### 2023-01-15 tensor_clustering/ieee_revision/response_s2.tex

%   The right hand side corresponds to the following hypothesis test for vectors: for an arbitrary $i \in T_1^c$, 
%   \begin{equation}\label{eq:hypo}
%       H_0: \my = \mtheta_1^*(i) \mx_1^* + \me \quad \leftrightarrow \quad H_1: \my = \mtheta_1^*(i) \mx_2^* + \me,
%   \end{equation}
%   where $\my = [\mat_1(\tY)]_{i:}$, $\mx_a^* = [\mat_1( \tS^* \times_2 \mTheta^* \mM_2^* \times_3 \cdots \times_K \mTheta^* \mM_K^*)]_{a:}$, and $\me$ consists of i.i.d.\ $N(0,\sigma^2)$ entries. By Neyman-Pearson lemma, the least square estimator is the best test for above hypothesis testing, where $\hat z_{LS}(i) = \argmin_{a = 1,2} \onormSize{}{\my - \mtheta_1(i)^* \mx_a^*}_F^2$.

%   Note that 
%   \begin{align}
%       \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F &\leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*)  \\
%       & \leq \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F   \max_{k \in [K]/\{1\}, a \in [r]} \onormSize{}{\mtheta_{k, z^{*, -1}_k(a)}}_2^{K-1} \\
%       & \leq  \sigma r^{(K-1)/2} p^{-(K-1)/2} 2 c_4 p^{(K-1)/2} r^{-(K-1)/2}\\
%       & \leq 2 c_4 \sigma, \label{eq:bound}
%   \end{align}
%   where the second inequality follows Lemma 6, the third inequality follows from property (ii) and the boundedness constraint in $\tP_{\tS}$ such that $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq \onormSize{}{\mS_{1:}^*}_F + \onormSize{}{\mS_{2:}^*}_F \leq 2c_4$.

%   Hence, we have 
%   \begin{align}
%       & \inf_{\hat z_1 \in \tP_1} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, \mtheta_k^*, \tS^*] \} \\
%       & \quad = 2 \bbP[ \hat z_{LS}(i) = 1 |  z_1^*(i) = 2, \mtheta_k^*, \tS^* ] \\
%       & \quad  = 2 \bbP[ \onormSize{}{\my - \mtheta_1(i)^* \mx_1^*}_F^2 \leq \onormSize{}{\my - \mtheta_1(i)^* \mx_2^*}_F^2 | z_1^*(i) = 2, \mtheta_k^*, \tS^*   ] \\
%       & \quad =  2 \bbP[ N(0,1) \geq \mtheta_1(i)^* \onormSize{}{\mx_1^* - \mx_2^*}_F /\sigma ] \\
%       & \quad \geq 2 \bbP[ N(0,1) \geq 2 c_4 ]   \geq c, \label{eq:inf}
%   \end{align}
%   for some positive constant $c$, where the calculation for the last equation follows from the calculation in equation (102) in \cite{han2020exact}.

%   Plugging the inequality~\eqref{eq:inf} to inequality~\eqref{eq:inf_lower}, then, we have 
%   \begin{equation}
%       \liminf_{p \rightarrow \infty}  \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]  \geq \liminf_{p \rightarrow \infty} \frac{ cp }{r^3} \geq c,
%   \end{equation}
%   where the last inequality follows by the condition $r = o(p^{1/3})$. By the discrete nature of the misclassification error, we have 
%   \begin{equation}
%       \liminf_{p \rightarrow \infty} \inf_{\tS^* \in  \tP_{\tS} \cap \{\Delta_{\min}^2(\tS^*) \leq 2\}}  \inf_{\hat z_{\text{stat}} \in \tP_z}\sup_{ (z^*, \mtheta^*) \in \tP_{z, \mtheta}} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]  \geq 1. 
%   \end{equation}
% %   for any $r < p$ and $p$,  we have 
% %   \begin{equation}
% %         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq \frac{ cp }{r^3} > 0,
% %   \end{equation}
% %   which indicates 
% %   \begin{equation}
% %         \inf_{\hat z_1 \in \tP_z} \bbE \left[ p \ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right] \geq 1, \quad \text{for every } p
% %   \end{equation}
% %   by the discrete nature of $p\ell(\hat z, z^*)$.
%   \end{proof}

%   \begin{rmk}[Where the triviality comes from]
%   The proof of Lemma~\ref{lem:trivial} does not use the angle gap of $\tS^*$ but use the Euclidean gap of $\tS^*$ when bound the key quantity $ \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F$ in \eqref{eq:bound}. By the boundedness constraint in $\tP_{\tS}$, we have $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F  \leq 2 c_4$. Then we can construct a small $\mtheta^*(i)$ to eliminate the constant Euclidean gap and make the classification problem very hard.  That's why the minimax bound~\eqref{eq:minminmax} is trivial with respect to the angle gap.
%   \end{rmk}

%   \begin{rmk}[Strongest position of $\mtheta$]
%   The proof of Lemma~\ref{lem:trivial} also indicates the strongest position of $\mtheta^*$ should be under the sup. Note that the key quantity in \eqref{eq:bound} relies on the Euclidean gap of $\tS^*$; i.e., 
%   \begin{equation}
%      \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\min}(\mTheta_k^*\mM_k^*)   \leq  \onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F \leq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\max}(\mTheta_k^*\mM_k^*).
%   \end{equation}
%   We can not infer the Euclidean gap of $\tS^*$ from the angle gap of $\tS^*$. Therefore, we need to design a specific $\mtheta^*$ to eliminate the gap $\onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F$ to make a hard case to solve the hypothesis test~\eqref{eq:hypo}.

%   If we take $\mtheta$ under the inf, then we need to consider arbitrary combination of $\mtheta^*$ and $\tS^*$. Take $(\mtheta^*, \tS^*)$ such that $\onormSize{}{ \mtheta_1^*(i) \mx_1^*  - \mtheta_1^*(i) \mx_2^*}_F  \geq  \mtheta_1^*(i) \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F \prod_{k = 2}^K \lambda_{\min}(\mTheta_k^*\mM_k^*)   \geq c \sigma p^{(K-1)/2} r^{-(K-1)/2}$. Plugging into the inequality~\eqref{eq:inf}, we have 
%   $ \inf_{\hat z_1 \in \tP_1} \{ \bbP[\hat z_1(i) = 1| z_1^*(i) = 2, \mtheta_k^*, \tS^*] +   \bbP[\hat z_1(i) = 2| z_1^*(i) = 1, \mtheta_k^*, \tS^*] \} =  0$ and thus $\inf_{\hat z_1 \in \tP_z} \bbE \left[ p\ell(\hat z_1, z_1^*) | z^*_k, \mtheta^*_k, \tS^*  \right]  \geq 0$, which is meaningless because the misclassification error always $\geq 0$. 
%   \end{rmk}

%   \begin{rmk}[Comparison with TBM] In TBM, the statistical impossibility is not trivial. We will have following results:

%   Assume $ r \lesssim p^{1/3}$. If the signal exponent satisfies $\gamma < -(K-1)$, then, for any true core tensor $\tS^* \in \tP_{\tS}$ such that $\Delta_{\rm Euc}^2 (\tS^*) = p^{\gamma}$, no estimator $\hat z_{\text{stat}}$ achieves exact recovery in expectation under the worst case; i.e.,

%   \begin{equation}
%   \inf_{\tS^* \in \tP_{TBM, \tS} \cap \{\Delta_{\rm Euc}^2(\tS^*) = p^{\gamma} \} }  \inf_{\hat z_{\text{stat}} \in \tP_z }\sup_{ (z^*, \mtheta^*): (z^*, \tS^*, \mtheta^*) \in \tP_{TBM}(\gamma)} \bbE \left[ p\ell(\hat z_{\text{stat}}, z) \right]\geq 1.
% \end{equation}
% The prove of above result is exactly the Proof of Lemma~\ref{lem:trivial} after replacing $\mtheta^* = \mathbf{1}$. Particularly, the inequality~\eqref{eq:bound} becomes 
% \begin{equation}
%      \onormSize{}{  \mx_1^*  -  \mx_2^*}_F  \leq \onormSize{}{ \mS^*_{1:} - \mS^*_{2:} }_F p^{(K-1)/2} r^{-(K-1)/2},
% \end{equation}
% and thus we need the Euclidean gap $\Delta_{\rm Euc}^2 (\tS^*) \leq \sigma^2  p^{-(K-1)} r^{K-1}$ to have $ \onormSize{}{  \mx_1^*  -  \mx_2^*}_F  \leq \sigma$. That's the why the TBM minimax rate is not trivial.
%   \end{rmk}

\bibliography{tensor_wang}
\bibliographystyle{apalike}

\end{document}
